Measuring Perceptions and Preferences for Meritocracy

### 2021-01-18 docs/paper.Rmd

---
title: "Measuring Perceptions and Preferences for Meritocracy"
css: "custom.css"
linestretch: '1.5'
link-citations: yes
author: "Juan Carlos Castillo, Julio Iturra & Francisco Meneses <br><br>"
abstract: "Economic and social inequalities have generated growing concern and crises across contemporary societies. One of the mechanisms proposed by the social sciences to explain the maintenance of inequality is the belief in meritocracy, which would legitimize economic disparities based on differences in effort and talent. Despite its wide use as a concept, empirical research on meritocracy is relatively novel, and as it is based upon multiple approaches and techniques, its interpretation can be difficult and confusing. Adding to these difficulties, is the fact that many such studies have relied upon secondary data to operationalize meritocracy, with significant variation in the use and interpretation of survey items. Based on a review of studies that measure meritocracy, this article identifies a series of drawbacks and inconsistencies within and between studies regarding the conceptualization and indicators of meritocracy. Based on this critical analysis, we propose a scale named the *Perceptions and Preferences for Meritocracy Scale*, which we test with confirmatory factor analysis, using data from an online survey study (N=2,141). Our results support the proposed conceptual structure, which not only distinguishes between perceptions and preferences, but also between meritocratic and non-meritocratic dimensions. Our discussion highlights the importance of considering these different dimensions in order to advance the study of meritocracy."
output:
  bookdown::html_document2:
    number_sections: false
  bookdown::pdf_document2:
    template: null
    toc: false
linkcolor: blue
bibliography:
  - ../input/bib/EconomiaMoral.bib
  - ../input/bib/meritocracy.bib
  - ../input/bib/invariance.bib
editor_options:
  chunk_output_type: console
geometry: margin=0.78in
header-includes:
  # - \usepackage[spanish,es-tabla,es-nodecimaldot]{babel}
  - \usepackage{times}           # Times New Roman
  - \usepackage{caption}
  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
---

@@

Findings the measure of NBI for \textit{personal wherewithal} also provide evidence for the convergent validity of PPM-S. Results indicate a positive correlation of this measure with meritocratic preference (r=.392), suggesting that both variables capture similar concepts. The NBI's factor depicts a positive association with unmeritocratic preference (r=.208). This finding is consistent with literature, inasmuch as the belief for meritocray is associated with the justification of inequality [@mccoy_Priming_2007;@MadeiraPrimesConsequencesSystematic2019]. Furthermore, results of our confirmatory factor analysis indicate that meritocratic perception and unmeritocratic preferences shows a positive and weak correlation.

>>>>

---
title: "Measuring Perceptions and Preferences for Meritocracy"
css: "custom.css"
linestretch: '1.5'
link-citations: yes
author: "Juan Carlos Castillo, Julio Iturra & Francisco Meneses <br><br>"
abstract: "Economic and social inequalities have generated growing concern and crises across contemporary societies. One of the mechanisms proposed by the social sciences to explain the maintenance of inequality is the belief in meritocracy, which would legitimize economic disparities based on differences in effort and talent. Despite its wide use as a concept, empirical research on meritocracy is relatively novel, and as it is based upon multiple approaches and techniques, its interpretation can be difficult and confusing. Adding to these difficulties, is the fact that many such studies have relied upon secondary data to operationalize meritocracy, with significant variation in the use and interpretation of survey items. Based on a review of studies that measure meritocracy, this article identifies a series of drawbacks and inconsistencies within and between studies regarding the conceptualization and indicators of meritocracy. Based on this critical analysis, we propose a scale named the *Perceptions and Preferences for Meritocracy Scale*, which we test with confirmatory factor analysis, using data from an online survey study (N=2,141). Our results support the proposed conceptual structure, which not only distinguishes between perceptions and preferences, but also between meritocratic and non-meritocratic dimensions. Our discussion highlights the importance of considering these different dimensions in order to advance the study of meritocracy."
output:
  bookdown::pdf_document2:
    template: null
    toc: false
  bookdown::html_document2:
    number_sections: false
linkcolor: blue
bibliography:
  - ../input/bib/EconomiaMoral.bib
  - ../input/bib/meritocracy.bib
  - ../input/bib/invariance.bib
editor_options:
  chunk_output_type: console
geometry: margin=0.78in
header-includes:
  # - \usepackage[spanish,es-tabla,es-nodecimaldot]{babel}
  - \usepackage{times}           # Times New Roman
  - \usepackage{caption}
  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
---

@@

Findings the measure of NBI for \textit{personal wherewithal} also provide evidence for the convergent validity of PPM-S. Results indicate a positive correlation of this measure with meritocratic preference (r=.392), suggesting that both variables capture similar concepts. The NBI's factor depicts a positive association with unmeritocratic preference (r=.208). This finding is consistent with literature, inasmuch as the belief for meritocray is associated with the justification of inequality [@McCall2017Exposurerisinginequality;@MadeiraPrimesConsequencesSystematic2019]. Furthermore, results of our confirmatory factor analysis indicate that meritocratic perception and unmeritocratic preferences shows a positive and weak correlation.
